Minimal trap spaces of Logical models are maximal siphons of their Petri net encoding

### 2022-05-09 cmsb22.tex
Comments

\sylvain{TBD}

@@

% \vangiang{n = 431 comes from Table 1 of~\cite{aghamiri2020automated}. I have checked this model. It has only 180 nodes. I will check others models come from~\cite{aghamiri2020automated}.}
% \sylvain{that's actually another model (called RA-map or something in the supplementary material), I think the other models should be ok.}
% \vangiang{Oh. Thank you for confirming that.}

@@

To assess the efficiency of the proposed method, we compared it with the state-of-the-art method PyBoolNet~\cite{klarner2017pyboolnet}.
We used a set of real-world Boolean models lying in various scales collected from various bibliographic sources.
These models are quite big (in size), complex (i.e., having high average in-degree, which is related to the number of prime-implicants) and most of them have never been fully analyzed.
We then applied the proposed method and PyBoolNet to computing minimal trap spaces of these real-world models.
It is notable that unlike existing analysis shown in the literature, we did not fix specific values for source nodes (i.e., some node \(v\) such that \(f_v = v\)) in these models.
To solve the ASP problems, we used the same ASP solver CLINGO~\cite{DBLP:journals/aicom/GebserKKOSS11} and the same configuration as that used in PyBoolNet~\cite{klarner2015computing}.
Specifically, we used the configuration \texttt{--dom-pref=32, --heu=domain, --dom-mod=7} (subset maximality). We ran all the benchmarks on a virtual machine whose environment is CPU: Intel(R) Core(TM) i7-3630QM 2.40GHz x 4, Memory: 8 GB, Ubuntu 18.04.2 64 bit. Finally, we set a time limit as one hour for each model.

% \sylvain{the \texttt{dom-pref} thing is deprecated since 2014 but still used in PyBoolNet, I think we might give the more modern version I used in the Python script, but we'll see}
% \vangiang{I also found that you did not use the \texttt{dom-pref} thing in the Python script. I think we should use the same configuration as that used in PyBoolNet.}
% \sylvain{as said above, this is deprecated, and has been since August 2014 \url{https://github.com/potassco/clasp/blob/master/CHANGES}, so I'd advise against using it, it might break any-time.}
% \vangiang{I see. Let's use the more modern version.}

@@

\begin{table}[!htb]
  \caption{Timing comparisons between the proposed method and PyBoolNet on the PyBoolNet's repository.
    Columns "\(n\)", "\(e\)", and "\(s\)" denote the numbers of nodes, interactions, and source nodes of each model, respectively. Column "\(|M|\)" denotes the number of minimal trap spaces and for each method is given the computation time in seconds.
    "DNF" means that the method did not finish the computation (with a maximum of 1000 trap spaces) within the timeout of 2 minutes. Number in bold indicate a ratio greater than three.}
  \centering
  \label{tab:pyboolnet_repo}
  \begin{tabular}{lrrrrrr}
    \toprule
    model & $n$ & $e$ & $s$ & $|M|$ & PyBoolNet (s) & Trappist (s) \\ \midrule
    arellano\_rootstem & 9 &&& 4 & 0.04 & \textbf{0.31}\\
    calzone\_cellfate & 28 &&& 27 & 0.03 & 0.08\\
    dahlhaus\_neuroplastoma & 23 &&& 32 & 0.04 & 0.11\\
    davidich\_yeast & 10 &&& 12 & 0.03 & 0.08\\
    dinwoodie\_life & 15 &&& 7 & 0.03 & 0.08\\
    dinwoodie\_stomatal & 13 &&& 1 & 0.04 & 0.05\\
    faure\_cellcycle & 10 &&& 2 & 0.04 & 0.04\\
    grieco\_mapk & 53 &&& 18 & 0.04 & 0.08\\
    irons\_yeast & 18 &&& 1 & 0.05 & 0.04\\
    jaoude\_thdiff & 103 &&& 1000+ & \textbf{1.73} & 0.40\\
    klamt\_tcr & 40 &&& 8 & 0.06 & 0.04\\
    krumsiek\_myeloid & 11 &&& 6 & 0.03 & 0.03\\
    multivalued & 13 &&& 4 & 0.02 & 0.02\\
    n12c5 & 11 &&& 5 & \textbf{34.37} & 0.06\\
    n3s1c1a & 2 &&& 2 & 0.02 & 0.03\\
    n3s1c1b & 2 &&& 2 & 0.02 & 0.03\\
    n5s3 & 4 &&& 3 & 0.03 & 0.03\\
    n6s1c2 & 5 &&& 3 & 0.03 & 0.04\\
    n7s3 & 6 &&& 3 & 0.02 & 0.02\\
    raf & 3 &&& 2 & 0.03 & 0.02\\
    randomnet\_n15k3 & 15 &&& 3 & 0.04 & 0.05\\
    randomnet\_n7k3 & 7 &&& 10 & 0.03 & \textbf{0.19}\\
    remy\_tumorigenesis & 34 &&& 25 & \textbf{1.79} & 0.13\\
    saadatpour\_guardcell & 13 &&& 1 & 0.02 & 0.03\\
    selvaggio\_emt & 56 &&& 1000+ & \textbf{1.03} & 0.23\\
    tournier\_apoptosis & 12 &&& 3 & 0.05 & 0.06\\
    xiao\_wnt5a & 7 &&& 4 & 0.06 & 0.03\\
    zhang\_tlgl & 60 &&& 156 & 0.26 & 0.17\\
    zhang\_tlgl\_v2 & 60 &&& 258 & 0.19 & 0.17\\
    \bottomrule
  \end{tabular}

  \sylvain{I'm not sure that \(e\) and \(s\) are very relevant, moreover they are not readily available\dots}
\end{table}

@@

\sylvain{Some models seem to have a different number of trap spaces here than in the other table (like RA apoptosis), for others quite large time differences (ERBB, inflammatory bowel)\dots}
\vangiang{Regarding the time differences, the reason might be that my implementation uses an ASP encoding based on choice rules instead of disjunctive rules or constraint rules.}
\vangiang{Regarding the differences in the numbers of minimal trap spaces, I have also observed this phenomenon. For example, for the Colitis\_associated\_colon\_cancer, the number of minimal trap spaces should be 10 (I am sure that this number is correct). For the TLGLSurvival model, the number of minimal trap spaces should be 318. Since PyBoolNet and Trappist return the same results for these models, I do not think that there is a problem with Trappist. Please check the model files.}
\end{document}

>>>>

@@

@@

To assess the efficiency of the proposed method, we compared it with the state-of-the-art method PyBoolNet~\cite{klarner2017pyboolnet}.
We used a set of real-world Boolean models lying in various scales collected from various bibliographic sources.
These models are quite big (in size), complex (i.e., having high average in-degree, which is related to the number of prime-implicants) and most of them have never been fully analyzed.
We then applied the proposed method and PyBoolNet to computing minimal trap spaces of these real-world models.
It is notable that unlike existing analysis shown in the literature, we did not fix specific values for source nodes (i.e., some node \(v\) such that \(f_v = v\)) in these models.
To solve the ASP problems, we used the same ASP solver CLINGO~\cite{DBLP:journals/aicom/GebserKKOSS11} and the same configuration as that used in PyBoolNet~\cite{klarner2015computing}.
Specifically, we used the configuration \texttt{-heuristic=Domain -enum-mod=domRec -dom-mod=3} (subset maximality, equivalent to the deprecated \texttt{--dom-pref=32 --heu=domain --dom-mod=7} used by PyBoolNet). We ran all the benchmarks on a virtual machine whose environment is CPU: Intel(R) Core(TM) i7-3630QM 2.40GHz x 4, Memory: 8 GB, Ubuntu 18.04.2 64 bit. Finally, we set a time limit as one hour for each model.

@@

\begin{table}[!htb]
  \caption{Timing comparisons between the proposed method and PyBoolNet on the PyBoolNet's repository.
    Columns "\(n\)", "\(e\)", and "\(s\)" denote the numbers of nodes, interactions, and source nodes of each model, respectively. Column "\(|M|\)" denotes the number of minimal trap spaces and for each method is given the computation time in seconds.
    "DNF" means that the method did not finish the computation (with a maximum of 1000 trap spaces) within the timeout of 2 minutes. Number in bold indicate a ratio greater than three.}
  \centering
  \label{tab:pyboolnet_repo}
  \begin{tabular}{lrrrrrr}
    \toprule
    model & $n$ & $e$ & $s$ & $|M|$ & PyBoolNet (s) & Trappist (s) \\ \midrule
    arellano\_rootstem & 9 &&& 4 & 0.04 & \textbf{0.31}\\
    calzone\_cellfate & 28 &&& 27 & 0.03 & 0.08\\
    dahlhaus\_neuroplastoma & 23 &&& 32 & 0.04 & 0.11\\
    davidich\_yeast & 10 &&& 12 & 0.03 & 0.08\\
    dinwoodie\_life & 15 &&& 7 & 0.03 & 0.08\\
    dinwoodie\_stomatal & 13 &&& 1 & 0.04 & 0.05\\
    faure\_cellcycle & 10 &&& 2 & 0.04 & 0.04\\
    grieco\_mapk & 53 &&& 18 & 0.04 & 0.08\\
    irons\_yeast & 18 &&& 1 & 0.05 & 0.04\\
    jaoude\_thdiff & 103 &&& 1000+ & \textbf{1.73} & 0.40\\
    klamt\_tcr & 40 &&& 8 & 0.06 & 0.04\\
    krumsiek\_myeloid & 11 &&& 6 & 0.03 & 0.03\\
    multivalued & 13 &&& 4 & 0.02 & 0.02\\
    n12c5 & 11 &&& 5 & \textbf{34.37} & 0.06\\
    n3s1c1a & 2 &&& 2 & 0.02 & 0.03\\
    n3s1c1b & 2 &&& 2 & 0.02 & 0.03\\
    n5s3 & 4 &&& 3 & 0.03 & 0.03\\
    n6s1c2 & 5 &&& 3 & 0.03 & 0.04\\
    n7s3 & 6 &&& 3 & 0.02 & 0.02\\
    raf & 3 &&& 2 & 0.03 & 0.02\\
    randomnet\_n15k3 & 15 &&& 3 & 0.04 & 0.05\\
    randomnet\_n7k3 & 7 &&& 10 & 0.03 & \textbf{0.19}\\
    remy\_tumorigenesis & 34 &&& 25 & \textbf{1.79} & 0.13\\
    saadatpour\_guardcell & 13 &&& 1 & 0.02 & 0.03\\
    selvaggio\_emt & 56 &&& 1000+ & \textbf{1.03} & 0.23\\
    tournier\_apoptosis & 12 &&& 3 & 0.05 & 0.06\\
    xiao\_wnt5a & 7 &&& 4 & 0.06 & 0.03\\
    zhang\_tlgl & 60 &&& 156 & 0.26 & 0.17\\
    zhang\_tlgl\_v2 & 60 &&& 258 & 0.19 & 0.17\\
    \bottomrule
  \end{tabular}
\end{table}

  \sylvain{I'm not sure that \(e\) and \(s\) are very relevant, moreover they are not readily available\dots}

@@

\sylvain{Some models seem to have a different number of trap spaces here than in the other table (like RA apoptosis), for others quite large time differences (ERBB, inflammatory bowel)\dots}
\vangiang{Regarding the time differences, the reason might be that my implementation uses an ASP encoding based on choice rules instead of disjunctive rules or constraint rules.}
\sylvain{Maybe\dots I don't know ASP well enough to be sure. The fact that you go through a predicate \texttt{p} which adds one level of indirection might also have some effect, I'm not sure.}

\vangiang{Regarding the differences in the numbers of minimal trap spaces, I have also observed this phenomenon. For example, for the Colitis\_associated\_colon\_cancer, the number of minimal trap spaces should be 10 (I am sure that this number is correct). For the TLGLSurvival model, the number of minimal trap spaces should be 318. Since PyBoolNet and Trappist return the same results for these models, I do not think that there is a problem with Trappist. Please check the model files.}
\sylvain{There are not among the files I had to change for BioLQM, so I'm not sure what I can check. For TLGL it should be quite obvious since 3 and 318 are pretty far away, but\dots I don't see any thing obviously wrong in the files. The 3 minimal trap spaces found are 3 stable-states, all 0 or with a single 1. They are the same also found using the BDD of BioLQM, so\dots}
\end{document}
